Abstractions for Type-Level Programming

### 2022-04-19 preprint.tex
Do TODOs
Generalized singleton type

When inferring the result type of a |dependent| method, our system lifts the method's body into a generalized singleton type.
This lifting will be precise for the subset of expressions that is representable in types, and approximative for the rest.
When we encounter an unsupported construct, we compute its type using the default mode, yielding a type |T| which we then integrate in the lifted body as |_: T|.

@@

\noindent
We could have equivalently defined |getName| by omitting the |dependent| qualifier and writing its result type explicitly.
The difference between the two would only be matter of syntax.
% TODO: say something about the fact that generalized singleton types are not singletons...

@@

Scala's type system permits uncontrolled side effects in programs.
Given the absence of an effect system, result types of methods do not convey any information about the potential use of side effects in the method body.
The situation is analogous for |dependent| methods.
Since we uniformly approximate all side effects, we avoid the situation where a type refers to a value that may be modified during the program execution.
For instance, if |z| is a mutable integer variable, we will never introduce |z| in a singleton type.
However, we can still assign singleton types to expressions containing |z|, for example, we can type |Cons(z, Nil())| as |{ Cons(_: Int, Nil()) }|.
% TODO: More generally, when dependently typed functions call into a non-dependently typed one, ...

>>>>

When inferring the result type of a |dependent| method, our system lifts the method's body into a generalized singleton type.
This lifting will be precise for the subset of expressions that is representable in types, and approximative for the rest.
When we encounter an unsupported construct, we compute its type using the default mode, yielding a type |T| which we then integrate in the lifted body as |_: T|\footnotemark.
\footnotetext{%
When a generalized singleton type contains a component of form |_: T|, then that type is not singleton: it contains more than one value.
}

@@

\noindent
We could have equivalently defined |getName| by omitting the |dependent| qualifier and writing its result type explicitly.
The difference between the two would only be matter of syntax.

@@

Scala's type system permits uncontrolled side effects in programs.
Given the absence of an effect system, result types of methods do not convey any information about the potential use of side effects in the method body.
The situation is analogous for |dependent| methods.
Since we uniformly approximate all side effects, we avoid the situation where a type refers to a value that may be modified during the program execution.
For instance, if |z| is a mutable integer variable, we will never introduce |z| in a singleton type.
However, we can still assign singleton types to expressions containing |z|, for example, we can type |Cons(z, Nil())| as |{ Cons(_: Int, Nil()) }|.
More generally, when a dependently-typed function call into a non-dependently-typed one, we approximate the type of that call to |(_: T)|, where |T| is the  declared result type of the non-dependently-typed function.
